SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-08 docs/questioning.tex
add material in Section III

\begin{enumerate}
\item We provide an exploitation model that puts application attacks
  based on executable source code injection under a common
  perspective.
\item We categorize and analyze previously proposed defenses based on
  criteria that are considered critical when building protection
  mechanisms.
\item A worrying observation is that sometimes researchers test the
  diagnostic performance of their mechanisms in a poor manner. This
  turns out as very critical because some of these mechanisms can
  actually be bypassed.
\item Also worrying is that researchers do not make their mechanisms
  available to the community, thus no one can either use their
  mechanism, or replicate their experiments.
\item We discuss how our observations may be the main reasons why the
  mechanisms are not adopted in practice.
\item We discuss what are the necessary criteria for new mechanisms
  and how our work can help researchers to design new countermeasures.
\end{enumerate}

@@

@@

As in medical diagnostic tests, application attack defense mechanisms
must demonstrate the presence of an attack. This, however, does not
make an application attack defense mechanism immediately useful. In
order for a detection mechanism to be useful it must be accurate, easy to
use, and economical. The accuracy of a detection mechanism is gauged
with the following metrics~\cite{TDR2013,GFDLS06,A00}:
\begin{itemize}
\item {\bf Sensitivity}, the probability that an attack will be
  caught.
\item {\bf Specificity}, the probability that a normal interaction
  will test negative.
\item {\bf Positive Predictive Value} ({\sc ppv}), the probability that a
  reported attack is a real attack. It is the conditional probability
  that an event is an attack if the detection mechanism flags it as
  such. 
\item {\bf Negative Predictive Value} ({\sc npv}), the probability that if
  nothing is reported no attack has taken place. It is the conditional
  probability that an event is not an attack given that the detection
  mechanism flags it as normal.
\end{itemize}

@@

Note that the calculations are carried out vertically in
Table~\ref{tab:sensitivity-specificity}; the ratio of attack events to
normal events does not enter into sensitivity and specificity. This is
what allows us to derive them based on test data, without using data
from a real production environment. At the same time, however, this is
what limits their usefulness. When we obtain a test result in a
production environment, be it negative or positive, we really want to
know how much we should be worried, or relaxed. In other words, if an
attack is detected in a production environment, how much should we be
worried? If no attack is detected in a production environment, how
relaxed should we be that no attack has indeed taken place?

@@

@@

In this section we categorize and describe the various mechanisms
developed to detect the attacks described in
Section~\ref{sec:attacks}. We accompany our descriptions with comments
related to the following dimensions: a) diagnostic performance, b)
computational performance, c) security, d) ease of use, and e)
detection locus. We treat f) availability separately in the
``Discussion" section (\ref{sec:discussion}).

>>>>

\begin{enumerate}
\item We provide an exploitation model that puts application attacks
  based on executable source code injection under a common
  perspective.
\item We categorize and analyze proposed defenses using a set of
  criteria that are important for building protection mechanisms.
\item We find that researchers often test the diagnostic performance
  of their mechanisms poorly. 
\item We also find that researchers do not often make their mechanisms
  available to the community, thus no one can either use their
  mechanism, or replicate their experiments.
\item We discuss how our observations may be the main reasons why the
  mechanisms are not adopted in practice.
\item We discuss what are the necessary criteria for new mechanisms
  and how our work can help researchers to design new countermeasures.
\end{enumerate}

@@

Research on application attack defense mechanisms has a dual
character. First, a defense mechanism may be important, and therefore
publishable, because it shows that an attack can be detected in a
reliable manner. Secondly, a defense mechanism may be important not
just as a research contribution, but as a practical tool, if it can be
put to use by administrators and users to shield their applications
and services against the detected attacks.

The detection of attacks in a reliable manner can be analyzed using
criteria common with other research fields:
\begin{itemize}
\item Statistical measures that show how reliable the detection really
  is.
\item Research practices that promote replication of the findings.
\end{itemize}

Whether a reliable attack defense mechanism has value in a practical
setting rests on a different set of criteria:
\begin{itemize}
\item What are the overheads imposed by the mechanism?
\item How easy is it to deploy and use the mechanism?
\item How robust is it against ways to circumvent it?
\item At which point of the computation flow does it detect an attack?
\end{itemize}

The first two criteria are \emph{Diagnostic Performance} and
\emph{Availability}. The other four criteria are \emph{Computational
  Performance}, \emph{Ease of Use}, \emph{Security}, and
\emph{Detection Locus}.

@@

As in medical diagnostic tests, application attack defense mechanisms
must demonstrate the presence of an attack. This, however, does not
make an application attack defense mechanism immediately useful. A
detection mechanism must be reliable. The accuracy of a detection
mechanism is gauged with the following
metrics~\cite{TDR2013,GFDLS06,A00}:
\begin{itemize}
\item {\bf Sensitivity}, the probability that an attack will be
  caught.
\item {\bf Specificity}, the probability that a normal interaction
  will test negative.
\item {\bf Positive Predictive Value} ({\sc ppv}), the probability that a
  reported attack is a real attack. It is the conditional probability
  that an event is an attack if the detection mechanism flags it as
  such. 
\item {\bf Negative Predictive Value} ({\sc npv}), the probability that if
  nothing is reported no attack has taken place. It is the conditional
  probability that an event is not an attack given that the detection
  mechanism flags it as normal.
\end{itemize}

@@

Note that the calculations are carried out vertically in
Table~\ref{tab:sensitivity-specificity}; the ratio of attack events to
normal events does not enter into sensitivity and specificity. This is
what allows us to derive them based on test data, without using data
from a real production environment. At the same time, this is what
limits their usefulness. When we obtain a test result in a production
environment, be it negative or positive, we really want to know how
much we should be worried, or relaxed. In other words, if an attack is
detected in a production environment, how much should we be worried?
If no attack is detected in a production environment, how relaxed
should we be that no attack has indeed taken place?

@@

All detection mechanisms extract a cost for their use, as they
introduce some amount of extra computation on an existing application.
The extra computation depends on the exact defense mechanism. For
example, it may be some form of run-time checking, or some form of
obfuscation. Also depending on the mechanism, the cost may be incurred
on different places: it may take place on a server, affecting service
performance, or it may take place on the client, or both. The
usefulness of a detection mechanism depends therefore on the
computational cost it requires and on where it extracts it, as
different overheads may be acceptable server-side than client-side.
For example, reporting absolute measurements gives little information
on the real overhead, unless separate measurements are given for the
system under study with and without the proposed mechanism. Percentage
measurements are normally better. Outliers may be important. It is
always necessary to describe the experimental setup.

Performance evaluation rests on strong foundations~\cite{jain1991} and
is a vibrant field as new technologies emerge~\cite{gregg2014}.
Although it may not be necessary to conduct a comprehensive
performance evaluation analysis for a defense mechanism, the more
performance evaluation is provided for a mechanism the more valuable
it becomes as a practical approach.

@@

We categorize and describe the various mechanisms developed to detect
the attacks described in Section~\ref{sec:attacks} on the dimensions
we proposed apart from availability, which we treat separately in the
``Discussion" section (\ref{sec:discussion}).
